---
title: "Paula 16-plex"
author: "Pat"
date: "22/11/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r}

library(rmarkdown)

```
#### set messages to FALSE on everything (prevents certain boring things from being shown in the results)
```{r}

knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,collapse = TRUE) 

```
#### PACKAGES
```{r}

library(reshape2)
library(ggplot2)
library(dplyr)
library(plotly)
library(viridis)
library(data.table)
library(pheatmap)
library(tidyverse)
library(ggthemes)
library(clipr)
library(tidyr)
library(matrixStats)
library(ggrepel)
library(cowplot)

```
## plus genefilter if needed
```{r}

##if (!requireNamespace("BiocManager", quietly = TRUE))
    ##install.packages("BiocManager")
##BiocManager::install("genefilter")
library(genefilter)

```
#### COLORS
```{r}

mycolors<-c(viridis(15))
felix_cols<-mycolors[c(5,2)]
felix_4cols<-mycolors[c(15,10,8,2)]
plain_cols1<-c("blue","gray")
plain_cols2<-c("gray","#481B6DFF")
pats_cols<-colorRampPalette(c("#FDE725FF", "white","#440154FF"))(21)
leos_cols<-colorRampPalette(c("white","blue"))(10)
leos_8cols<-mycolors[c(5,5,5,5,2,2,2,2)]
mukha_cols<-mycolors[c(5,5,5,5,5,2,2,2,2,2,4,4,4,3,3,3)]
beaus_colors<-colorRampPalette(c("white","#28aee4","#081534"))(25)

```
#### LOAD, NORMALIZE AND ORGANIZA DATA (protein data b is the failed labeling for the NAD)
```{r}

## load the dataset
protein_data<-read_csv(file="protein_quant_16_plex.csv")
##protein_data<-read_csv(file="protein_data.csv")

## sums with all data
#sums<-protein_data %>% 
 #select(Ctrl_1:GAA_5) %>%
  #colSums();correction_factors<-max(sums)/sums;correction_factors

## creatine 2 and ctrl  are way too far off

##make column sums
sums<-protein_data %>% 
  select(MCF_ctrl_1:MDA_KD_4) %>%
  colSums()
##correction factors from the column sums
correction_factors<-max(sums)/sums

correction_factors


##corrected data
protein_data_norm<-protein_data %>% 
  select(MCF_ctrl_1:MDA_KD_4) %>% 
  sweep(2, correction_factors, "*")

sums<-protein_data_norm %>% 
  select(MCF_ctrl_1:MDA_KD_4) %>%
  colSums()

##corrected data


##bind the normalized data with the rest of the dataset
n_protein_data<-protein_data %>% 
  select(-c(MCF_ctrl_1:MDA_KD_4)) %>% 
  cbind(protein_data_norm) 



## make some new columns that store the mean values 
n_protein_data<-n_protein_data %>% 
  mutate(MCF_log_KD_ctrl=log2((MCF_KD_1+MCF_KD_2+MCF_KD_3+MCF_KD_4)/(MCF_ctrl_1+MCF_ctrl_2+MCF_ctrl_3+MCF_ctrl_4))) %>% 
  mutate(MDA_log_KD_ctrl=log2((MDA_KD_1+MDA_KD_2+MDA_KD_3+MDA_KD_4)/(MDA_ctrl_1+MDA_ctrl_2+MDA_ctrl_3+MDA_ctrl_4)))


```
#### PVALUE STATS
```{r}

## use the genefilter package to calculate the p-values
f1=factor(c(0,0,0,0,1,1,1,1)) ## need these factors for the different ttest



Stats_MCF<-n_protein_data %>% select(MCF_ctrl_1:MCF_KD_4) %>% 
  as.matrix() %>% 
  rowttests(f1) %>% ## performs an ttest
  select(p.value)

Stats_MDA<-n_protein_data %>% select(MDA_ctrl_1:MDA_KD_4) %>% 
  as.matrix() %>% 
  rowttests(f1) %>% ## performs an ttest
  select(p.value)



```

#### COMBINE STATS WITH DATA
```{r}
low_cutoff<--0.5

## MCF-7

n_protein_data_stats<-n_protein_data %>% 
  cbind(Stats_MCF) %>% ## takes the pvalues and binds them to the dataset a new row
  rename(MCF_pval=p.value) %>% 
  mutate(MCF_adj_p=p.adjust(MCF_pval,method="BH")) %>% # adjust the p values for multiple-hypothesis testing
  mutate(MCF_log_adj_p=-log10(MCF_adj_p)) %>% 
  mutate(MCF_log_p=-log10(MCF_pval)) %>% 
  mutate(MCF_significant=case_when(MCF_log_p>1.3 & (MCF_log_KD_ctrl<low_cutoff | MCF_log_KD_ctrl>0.5) ~ "significant"))


n_protein_data_stats %>% group_by(MCF_significant) %>% 
  summarise(n=n())

## MDA 231

n_protein_data_stats<-n_protein_data_stats %>% 
  cbind(Stats_MDA) %>% ## takes the pvalues and binds them to the dataset a new row
  rename(MDA_pval=p.value) %>% 
  mutate(MDA_adj_p=p.adjust(MDA_pval,method="BH")) %>% # adjust the p values for multiple-hypothesis testing
  mutate(MDA_log_adj_p=-log10(MDA_adj_p)) %>% 
  mutate(MDA_log_p=-log10(MDA_pval)) %>% 
  mutate(MDA_significant=case_when(MDA_log_p>1.3 & (MDA_log_KD_ctrl<low_cutoff | MDA_log_KD_ctrl>0.5) ~ "significant"))


n_protein_data_stats %>% group_by(MDA_significant) %>% 
  summarise(n=n())


```
#### ORGANIZE THE DATA PVALUES
```{r}

##n_protein_data_stats_sig<-n_protein_data_stats %>% 
  ##mutate(significant_A_PISA=ifelse(pval_A_PISA<0.05, "significant","not significant"),
        # significant_B_PISA=ifelse(pval_B_PISA<0.05, "significant","not significant")) ## see significance

##n_protein_data_stats_sig %>% group_by(significant_A_PISA) %>% 
  ##summarise(proteins=n())

##n_protein_data_stats_sig %>% group_by(significant_B_PISA) %>% 
  ##summarise(proteins=n())

##export the data from r
##write.csv(n_protein_data_stats_sig,"Ecoli_protein_data_stats_sig.csv")

```
#### ORGANIZE THE DATA ADJUSTED PVALUES
```{r}





```




### VOLCANO PLOT WITH NON-ADJUSTED p-values
```{r,fig.width=10,fig.height=7}

## make a volcano MCF-7
VP_PISA_MCF<-n_protein_data_stats %>%
  arrange(desc(MCF_significant)) %>% 
  ggplot(aes(x=MCF_log_KD_ctrl,
             y=MCF_log_p,
             description=`Gene`,
             color=MCF_significant))+
  geom_point(alpha=0.7,size=1)+
  theme_light()+
  scale_colour_manual(values=plain_cols1)+
  theme(axis.text=element_text(size=12))

##VP_PISA_MCF
ggplotly(VP_PISA_MCF)


## make a volcano MDA231
VP_PISA_MDA<-n_protein_data_stats %>%
  arrange(desc(MDA_significant)) %>% 
  ggplot(aes(x=MDA_log_KD_ctrl,
             y=MDA_log_p,
             description=`Gene`,
             color=MDA_significant))+
  geom_point(alpha=0.7)+
  theme_light()+
  scale_colour_manual(values=plain_cols1)+
  theme(axis.text=element_text(size=12))

##VP_PISA_MDA
ggplotly(VP_PISA_MDA)


```

## plot specific examples (based on the ttest showing differences between the fractions)
## Top 20 only

```{r}
##filter(Gene=="SPR") %>%
##slice_min(adj_pval_Anova,n=40)## us in place of filter to get the top anova-significant

plot_specific_examples<-n_protein_data_stats %>%
  filter(Gene=="AGO2") %>% 
  slice_min(MCF_pval,n=10) %>% 
   select(Gene,MCF_ctrl_1:MDA_KD_4) %>% 
   ## this is the filter function, which we set up to filter by gene. If  you replace this with slicn_min, it will instead pull out the lowest Anova pvalues
  pivot_longer(!Gene,names_to = "sample", values_to="Intensity") %>% 
  separate(sample, into=c("cell_line","treatment","replicate"),sep="_",) %>% 
  ggplot(aes(x=treatment,y=Intensity))+
  facet_wrap(cell_line~Gene,scales="free_y")+
  geom_boxplot(position="dodge")+
  geom_jitter(width=0.2)+
  theme(axis.text.x = element_text(angle = 90))+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  theme_bw()+
  expand_limits(y=0)
    

plot_specific_examples

```

plot examples of several proteins if interest



```{r}

of_interest<-c("SORCS2")

plot_specific_examples<-n_protein_data_stats %>%
  filter(Gene %in% of_interest) %>% 
   select(Gene,MCF_ctrl_1:MDA_KD_4) %>% 
   ## this is the filter function, which we set up to filter by gene. If  you replace this with slicn_min, it will instead pull out the lowest Anova pvalues
  pivot_longer(!Gene,names_to = "sample", values_to="Intensity")  %>% 
  separate(sample, into=c("cell_line","treatment","replicate"),sep="_",) %>% 
  ggplot(aes(x=treatment,y=Intensity))+
  facet_wrap(cell_line~Gene)+
  geom_boxplot(position="dodge")+
  geom_jitter(width=0.2)+
  theme(axis.text.x = element_text(angle = 90))+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  theme_bw()+
  expand_limits(y=0)

plot_specific_examples



```

Write out the fully processed data


```{r}
write_csv(n_protein_data_stats,file="processed_data.csv")


```


MA plot

```{r}

n_protein_data_stats_for_MA<-n_protein_data_stats %>%
  mutate(MCF_rowprod=MCF_ctrl_1*MCF_ctrl_2*MCF_ctrl_3*MCF_ctrl_4*
           MCF_KD_1*MCF_KD_2*MCF_KD_3*MCF_KD_4,
         MDA_rowprod=MDA_ctrl_1*MDA_ctrl_2*MDA_ctrl_3*MDA_ctrl_4*
           MDA_KD_1*MDA_KD_2*MDA_KD_3*MDA_KD_4) %>% 
  mutate(MCF_A=(log2(MCF_rowprod)/2)) %>% 
  mutate(MDA_A=(log2(MDA_rowprod)/2))




```


```{r}

## make an MA plot of MCF-7
VP_PISA_MCF<-n_protein_data_stats_for_MA %>%
  arrange(MCF_significant) %>% 
  ggplot(aes(y=MCF_log_KD_ctrl,
             x=MCF_A,
             description=`Gene`,
             color=MCF_significant))+
  geom_point(alpha=0.7,size=1)+
  theme_light()+
  scale_colour_manual(values=plain_cols1)+
  theme(axis.text=element_text(size=12))

##VP_PISA_MCF
ggplotly(VP_PISA_MCF)



## make an MA plot of MDA 231
VP_PISA_MCF<-n_protein_data_stats_for_MA %>%
  arrange(MCF_significant) %>% 
  ggplot(aes(y=MDA_log_KD_ctrl,
             x=MDA_A,
             description=`Gene`,
             color=MDA_significant))+
  geom_point(alpha=0.7,size=1)+
  theme_light()+
  scale_colour_manual(values=plain_cols1)+
  theme(axis.text=element_text(size=12))

##VP_PISA_MCF
ggplotly(VP_PISA_MCF)






```


